GravExplain: Demonstrating continuous gravitational wave searches using sound and a table-top interferometer

### 2020-05-05 paper/ifo-complexAudio.tex
some text editing and schematic editing

A natural succession to observing simple tones is to use complex audio signal, such as music and speech. 
The Viterbi algorithm and tracking more broadly is not applicable to these signals---music and speech are constantly changing over a broad spectrum. 
Instead of tracking a changing frequency over time, here we are interested in being able to record the interference pattern and play back these complex audio signal. 
We use the interferometer as an `optical microphone', using light to capture sound. 
The device has precedence in the laser microphones~\cite{laser_microphone} used in the defence industry which operate on a variety of related principles.

@@

\subsection{The photodiode method}
\label{sec:photodiode}
% advanced method: explain how we capture data

For the optical microphone to be sensitive to speech and music it needs a sampling rate far higher than the $30\,{\rm Hz}$ of the webcam used in Sections~\ref{sec:single_tone} and ~\ref{sec:viterbi_wandering}. 
Audible frequencies are in the range $\sim 20\,{\rm Hz}$--$20\,{\rm kHz}$. 
Speech intelligibility requires frequencies up to $3\,{\rm kHz}$ and music requires up to and beyond $8\,{\rm kHz}$~\cite{speech_intelligibility}. 
Therefore the optical microphone data requires a sample rate of at least $16\,{\rm kHz}$ to capture both speach and music. 
We achieved this with a photodiode.

A photodiode is an electrical component that acts as a regular diode wheen no light is incident on it, blocking any current flow in the `reverse' direction.
As the intensity of incident light rises, it becomes increasingly conductive in the reverse direction. 
% Fundamentally, a current is created in the photodiode by the photo-electric effect.
We place a (OSRAM BPW21) photodiod in reverse-bias over an (LM358) op-amp. 
This creates a photo-detector that produces a voltage dependent on the incident intensity. 
The photodiode recored the interference pattern at roughly the same off-centre position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} (however we saw no difference when monitoring at other positions in the interference pattern).
The photodiode was mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath.

The voltage signal from the photo-detector is captured by a (MCP3008) $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2. 
Together, these sampled the signal at roughly $16\,{\rm kHz}$, our desired limit.

@@

\begin{figure}
	\includegraphics[width=0.49\textwidth]{figures/ifo_schematic_photodiode.pdf}
	\caption{Michelson interferometer schematic with the intensity of the interference pattern at a point measured by a photodiode. See Figure~\ref{fig:ifo_schematic_webcam} for details on the optical configuration.}
	\label{fig:ifo_schematic_podo}
\end{figure}

@@

\begin{figure}
	\includegraphics[width=0.49\textwidth]{figures/butterworth_150_3000-cropped.pdf}
	\caption{Butterworth bandpass filter frequency response, any amplitude beyond -3dB is significant attenuation (half power). The green vertical lines show the band limits of 150Hz and 3kHz. Note the flat response within the band.}
	\label{fig:butterworth}
\end{figure}

>>>>

A natural succession to observing simple tones is to use complex audio signal, such as music and speech. 
The Viterbi algorithm and tracking more broadly is not applicable to these signals---music and speech are constantly changing over a broad spectrum. 
Instead of tracking a changing frequency over time, here we are interested in being able to record the interference pattern and play back these complex audio signals. 
We use the interferometer as an `optical microphone', using light to capture sound. 
The device has precedence in the laser microphones~\cite{laser_microphone} used in the defence industry which operate on a variety of related principles.

@@

\subsection{Recording data via photodiode}
\label{sec:photodiode}
% advanced method: explain how we capture data

Speech and music recordings require a higher sample rate than that of the webcam ($30\,{\rm Hz}$) used in Sections~\ref{sec:single_tone} and ~\ref{sec:viterbi_wandering}. 
Audible frequencies are in the range $\sim 20\,{\rm Hz}$--$20\,{\rm kHz}$. 
Speech intelligibility requires frequencies up to $3\,{\rm kHz}$ and music requires up to and beyond $8\,{\rm kHz}$~\cite{speech_intelligibility}. 
Therefore the optical microphone data requires a sample rate of at least $16\,{\rm kHz}$ to capture both speech and music. 
We achieved this with a photodiode.

A photodiode is an electrical component that acts as a regular diode wheen no light is incident on it, blocking any current flow in the `reverse' direction.
As the intensity of incident light rises, it becomes increasingly conductive in the reverse direction. 
% Fundamentally, a current is created in the photodiode by the photo-electric effect.
We place a (OSRAM BPW21) photodiod in reverse-bias over an (LM358) op-amp. 
This creates a photo-detector that produces a voltage dependent on the incident intensity. 
The photodiode records the interference pattern at roughly the same off-centre position as the webcam in Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} (however we see no difference when monitoring at other positions in the interference pattern).
The photodiode was mounted on a cloth screen re-purposed from the dismantled commercial speaker, with the electrical leads connected underneath.

The voltage signal from the photo-detector is captured by a (MCP3008) $10$-bit analog-to-digital converter (ADC) connected to a Raspberry Pi Model 3 v1.2. 
Together, these sampled the signal at roughly $16\,{\rm kHz}$.

@@

\begin{figure}
	\includegraphics[width=0.49\textwidth]{figures/ifo_schematic_photodiode_edit.pdf}
	\caption{
Michelson interferometer schematic. 
The schematic is identical to Figure~\ref{fig:interference_pattern} except that the data is now recorded using a photodiode and Raspberry Pi instead of a webcam. 
}
	\label{fig:ifo_schematic_podo}
\end{figure}

@@

\begin{figure}
	\includegraphics[width=0.49\textwidth]{figures/butterworth_150_3000-cropped.pdf}
	\caption{Butterworth bandpass filter frequency response, any amplitude beyond $-3\,{\rm dB}$ is significant attenuation (half power). The green vertical lines show the band limits of $150\,{\rm Hz}$ and $3\,{\rm kHz}$. Note the flat response within the band.}
	\label{fig:butterworth}
\end{figure}

### 2020-05-05 paper/ifo-introduction.tex
some text editing and schematic editing

% general gw intro
In 2015, gravitational waves were observed for the first time from the merger of two black holes in a binary system~\cite{GW150914}. 
The observation, made by the Laser Interferometer Gravitational-wave Observatory~\citep[LIGO]{AdvancedLIGO:2015}, marked a breakthrough in modern astrophysics and revealed a new means to observe the universe. 
Since 2015, the LIGO and Virgo~\cite{AdvancedVirgo:2015} observatories have made numerous detections of binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi,GW190425} mergers. 
In recent years, there has been increased public interest in gravitational wave science. 
Many gravitational wave research groups around the world have produced demonstrations and activities to explain the topic to a general audience.
Activities range from hands-on demonstrations, exhibitions~\cite{TTExhibit:2020,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, and online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, to phone apps~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online}%\han{look up Sarah Farmer's project}.

% what are gws and how do we detect them...
Gravitational waves are a prediction of Einstein's theory of general relativity. 
They are disturbances in space-time caused by the acceleration of massive objects. 
The effect of gravitational waves is a change in lengths; a `stretching and squeezing' of the distance between objects.
Observatories such as LIGO, Virgo, and KAGRA~\cite{KAGRA:2013} are laser interferometers; they use the interference of laser light to measure changes in distance. 
%As shown in an aerial view of LIGO Hanford in Figure~\ref{fig:ligo_pic}, they use beam arms kilometres long.
These observatories are extremely complex, but are fundamentally are based on the Michelson interferometer. 
Table-top Michelson interferometers are commonly used in undergraduate lab experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.

% continuous gws
To date the network of gravitational wave observatories have observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}. 
However, the network is also searching for continuous gravitational waves; persistent periodic signals at near-monochromatic frequencies.
Continuous waves may be emitted at constant frequencies, or may wander slowly in frequency over time. 
One of the prime targets of continuous wave searches are rotating neutron stars in low mass X-ray binaries (LMXBs).
LMXBs are binary systems containing a compact object (such as a neutron star or a black hole) in orbit with a low mass stellar companion~\cite{xraybinaries:1997}. 
The LMXB Scorpius X-1 is an LMXB which is particularly bright in X-rays. 
It is a prime target for continuous wave searches from LMXBs and numerous, as yet unsuccessful, searches have been performed~\cite{ScoX1O2Viterbi:2019,RadiometerO1O2:2019,SearchCrossCorrO1:2017}.

% sound analogy
Here, we demonstrate the use of a Michelson interferometer as a tool to explain the search for continuous waves to a general audience.
The analogy of sound is commonly used in explaining gravitational wave observations. 
%; the analogy of listening to the universe is often used to explain gravitational wave observations.
%; transient gravitational wave signals have been converted into audio.
In this work we utilise this analogy by playing sound close to a Michelson interferometer, simulating a gravitational-wave signal. 
Audio of constant or slowly wandering frequency tones provide an analogue of a continuous wave signal.
Continuous wave data analysis techniques are also demonstrated. 
The techniques used in this work are similar to those used to search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}, particularly for the wandering frequency demonstration.

This demonstration can be used to explain the search for continuous waves to a general audience as well as undergraduate students. 
The analysis methods applied here follow signal processing techniques used in some of the most recent searches for continuous waves~\cite{ScoX1O2Viterbi:2019}. 
As an undergraduate lab experiment, this demonstration covers topics of interest to both physics and engineering students.

In this work we detail our table-top interferometer design in Section~\ref{sec:ifo} and describe observing a single note from a speaker in Section~\ref{sec:single_tone}. In Section~\ref{sec:viterbi_wandering}, we demonstrate the analysis techniques for observing a wandering frequency signal.
In Section~\ref{sec:optical_microphone}, we demonstrate capturing and playing back complex audio, such as music and speech. This demonstration of an `optical microphone' serves as a more general exhibition of signal processing. We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}.

>>>>

% general gw intro
In 2015, gravitational waves were observed for the first time from the merger of two black holes in a binary system~\cite{GW150914}. 
The observation, made by the Laser Interferometer Gravitational-wave Observatory~\citep[LIGO]{AdvancedLIGO:2015}, marked a breakthrough in modern astrophysics and revealed a new means to observe the universe. 
Since 2015, the LIGO and Virgo~\cite{AdvancedVirgo:2015} observatories have made numerous detections of binary black hole~\cite{GW151226,GW170104,GW170814} and binary neutron star~\cite{GW170817,GW170817multi,GW190425} mergers. 
Gravitational-wave detection has led to increased public and media interest in gravitational-wave science. 
Gravitational-wave research groups around the world have produced demonstrations and activities to explain their research to general audiences. 
Activities range from hands-on demonstrations, exhibitions~\cite{TTExhibit:2020,LIGOScienceEducationCentre:online,GravityDiscoveryCentre:online}, and online data analysis tutorials~\cite{GWOSC:online,LOSC:2015}, to phone apps~\cite{LaserLabs:online,SciVR:online,chirp:online}, online games~\cite{BlackHoleHunter:online}, and musical performances~\cite{ArthurJeffesMusic:online,GravitySynthLeonTrimble:online}%\han{look up Sarah Farmer's project}.

% what are gws and how do we detect them...
Gravitational waves are a prediction of Albert Einstein's theory of general relativity~\cite{Einstein:1916}. 
They are disturbances in spacetime caused by the acceleration of massive objects. 
The effect of gravitational waves is a change in lengths: a `stretching and squeezing' of the distance between objects.
Observatories such as LIGO, Virgo, and KAGRA~\cite{KAGRA:2013} are laser interferometers, they use the interference of laser light to measure changes in distance. 
%As shown in an aerial view of LIGO Hanford in Figure~\ref{fig:ligo_pic}, they use beam arms kilometres long.
Gravitational-wave observatories are extremely complex, but are fundamentally are based on the Michelson interferometer. 
Table-top Michelson interferometers are commonly used in undergraduate lab experiments~\cite{UgoliniEtAl:2019} and to demonstrate the science of gravitational wave detection to non-specialist audiences~\cite{ThorLabsIFO,NikhefIFO,TTExhibit:2020,LIGOIFOGlue,LIGOIFOMagnets}.

% continuous gws
To date the network of gravitational wave observatories have observed short-duration transient signals~\cite{GWTC-1:2018,GWOSC:online}. 
However, the network is also searching for continuous gravitational waves; persistent periodic signals at near-monochromatic frequencies.
Continuous waves may be emitted at constant frequencies, or may wander slowly in frequency over time. 
Rotating neutron stars in low mass X-ray binaries (LMXBs) are one of the targets for continuous-wave searches. 
LMXBs are binary systems containing a compact object (such as a neutron star or a black hole) in orbit with a low mass stellar companion~\cite{xraybinaries:1997}. 
Scorpius X-1 is an LMXB which is particularly bright in X-rays. 
It is a prime target for continuous wave searches from LMXBs and numerous, as yet unsuccessful, searches have been performed~\cite{ScoX1O2Viterbi:2019,RadiometerO1O2:2019,SearchCrossCorrO1:2017}.

% sound analogy
We use the analogy of sound to demonstrate gravitational-wave science with a table-top experiment (see also Section~\ref{sec:ifo}). 
We use similar data analysis techniques to those used in the search for continuous waves in LIGO and Virgo data~\cite{ScoX1O2Viterbi:2019,ScoX1ViterbiO1:2017,SuvorovaEtAl:2017,SuvorovaEtAl:2017}.
As an undergraduate lab experiment this demonstration can be used to teach topics ranging from gravitational-wave detection and analysis to signal processing and speech enhancement. 
It can also be adapted for use as an outreach tool to explain gravitational-wave science to a general audience.

In this work we present a selection of demonstrations using different types of audio signals ranging from simple tones, analogous to continuous wave signals, to more complex audio of music and speech. 
In Section~\ref{sec:ifo} we detail our table-top interferometer design. 
We demonstrate observing a single note from a speaker in Section~\ref{sec:single_tone}.
In Section~\ref{sec:viterbi_wandering} we replace the single note with a slowly changing frequency and analyse the signal using techniques from continuous-wave searches. 
In Section~\ref{sec:optical_microphone}, we demonstrate capturing and playing back complex audio, such as music and speech. 
This demonstration of an `optical microphone' serves as a more general exhibition of signal processing. 
We suggest avenues of future work in Section~\ref{sec:future_work} and draw conclusions in Section~\ref{sec:conclusions}.

### 2020-05-05 paper/ifo-singleTone.tex
some text editing and schematic editing

The intensity of the interference pattern is measured at a single point on the screen using a commercial USB webcam placed to view the screen (see Figure~\ref{fig:ifo_schematic_webcam}). 
The webcam's sample rate is $30\,{\rm Hz}$, which limits the spectral content of observable signals to below $15\,{\rm Hz}$, the Nyquist frequency.
A tone under $15\,{\rm Hz}$ is played through the speaker for a minute and the interference pattern recorded with the webcam.

The resulting data is a video capture of the whole interference pattern. 
As only the frequency of the interference ring motion is needed, we take data from a single off-centre pixel indicated by the red dot in Figure~\ref{fig:interference_pattern}. 
We take the green-channel of the video as an approximation to the total intensity as the laser produces monochromatic green light (at $532\,{\rm nm}$).

We played this time series of intensity over time as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
The two tones played together sound similar and don’t produce any noticeable beating.
To quantify the precision, we recorded a signal from an injected tone of $2.09\,{\rm Hz}$. 
From the Fourier transform shown in Figure~\ref{fig:webcam_spectrum}, we measured a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$, a result within two standard deviations. 
The experiment is able to recover a single tone confidently. 
%This showed that the experiment observes a single tone confidently.

\begin{figure}
	\includegraphics[width=.49\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
	\caption{Fourier amplitudes of intensity time series of interference pattern as captured by a webcam for an injected tone of $2.09\,{\rm Hz}$. The peak (with amplitude $17.5$) was measured at $(2.099\pm 0.008) \mathrm{Hz}$. The actual maximum is achieved at $0\,{\rm Hz}$, but corresponds to the signal having a non-zero average value, so we don’t display or consider it.}
	\label{fig:webcam_spectrum}
\end{figure}

>>>>

The intensity of the interference pattern is measured at a single point on the screen using a commercial USB webcam placed to view the screen (see Fig.~\ref{fig:ifo_schematic_webcam}). 
The webcam's sample rate is $30\,{\rm Hz}$, which limits the spectral content of observable signals to below $15\,{\rm Hz}$, the Nyquist frequency.
A tone under $15\,{\rm Hz}$ is played through the speaker for a minute and the interference pattern recorded with the webcam.

The resulting data is a video capture of the whole interference pattern. 
As only the frequency of the interference ring motion is needed, we take data from a single off-centre pixel indicated by the red dot in Fig.~\ref{fig:interference_pattern}. 
We take the green-channel of the video as an approximation to the total intensity as the laser produces monochromatic green light (at $532\,{\rm nm}$).

As an initial inspection, we play the timeseries of the intensity using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
The two tones sound similar by ear and don't produce any noticeable beating. 
The result is quantified in Fig.~\ref{fig:webcam_spectrum}. 
A tone of $2.09\,{\rm Hz}$ is played through the speaker and the data recorded via webcam. 
Figure~\ref{fig:webcam_spectrum} shows the Fourier transform of the recording. 
We measure a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$ \han{[James: I added in that this is the $1\sigma$ uncertainty, is that correct?] (where the error represents the $1\sigma$ uncertainty).}
The demonstration is able to recover a single frequency tone to within two standard deviations. 
%We played this time series of intensity over time as an audio recording using the \textbf{io.wavfile.write} function from the Scipy~\cite{scipy} package in Python~\cite{python}. 
%The two tones played together sound similar and don’t produce any noticeable beating.
%To quantify the precision, we recorded a signal from an injected tone of $2.09\,{\rm Hz}$. 
%From the Fourier transform shown in Figure~\ref{fig:webcam_spectrum}, we measured a peak amplitude at $(2.099\pm 0.008)\,{\rm Hz}$, a result within two standard deviations. 
%The experiment is able to recover a single tone confidently. 
%This showed that the experiment observes a single tone confidently.

\begin{figure}
	\includegraphics[width=.49\textwidth]{figures/webcam_expt_4_0209-cropped.pdf}
	\caption{\label{fig:webcam_spectrum}
Single injected tone results. 
The Fourier amplitude of the intensity timeseres recorded by webcam. 
The frequency of the injection is $2.09\,{\rm Hz}$. 
The recovered frequency peak is at $(2.099\pm 0.008)\,{\rm Hz}$. 
%The peak (with amplitude $17.5$) has a frequency was measured at $(2.099\pm 0.008) \mathrm{Hz}$. The actual maximum is achieved at $0\,{\rm Hz}$, but corresponds to the signal having a non-zero average value, so we don’t display or consider it.
}

\end{figure}

### 2020-05-05 paper/ifo-tableTopGWs.tex
some text editing and schematic editing

The basic phyiscs of gravitational-wave detection can be demonstrated with a table-top Michelson iterferometer: a common optical configuration which can be used to measure changes in distance. 
A simple Michelson interferometer comprises of a laser, a beamsplitter, and two mirrors. 
Laser light is split evenly by a beamsplitter into two arms at right angles to each other. 
The beams travel to the end of each arm where they are reflected by mirrors. 
The beams then travel back to the beamsplitter where they recombine. 
The combination of the two beams produces an interference pattern which is dependent on the relative distance traveled by each of the beams. 
A change in arm length results in a change in the interference pattern.

Gravitational-wave detectors are large, complex interferometer.
The LIGO and Virgo detectors have $4\,{\rm km}$ and $3\,{\rm km}$ arms respectively.
They are sensitive to miniscule changes in distance, for example the first detection of a binary black hole merger had a peak gravitational-wave strain of $10^{-21}$~\cite{GW150914}.
However, their design is based on the Michelson interferometer. 
Detection and analysis concepts can be demonstrated with table-top optical equipment.

A commonly-used analogy in explaining gravitational-wave science is sound, although it is important to note that gravitational waves are not sound. 
Signals of binary black hole and neutron star mergers can be converted into audio as an aid to explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online}. 
Sound vibrations are a simple means to move the components of a demonstration interferometer. 
Audio files can be used as a simulated gravitational-wave signal in a table-top experiment. 
The vibrations move the mirrors, changing the length of the interferometer arms, and the resulting interference pattern. 
We use sound throughout this work as a means of demonstrating gravitaitonal-wave signals and the data analysis and signal processing techniques used to detect them.

\begin{figure}
	\includegraphics[width=0.49\textwidth]{figures/ifo_schematic_webcam.pdf}
	\caption{Michelson interferometer schematic with interference pattern recorded by a commercial webcam. The laser (top-left) is a Class 2 $532\,{\rm nm}$ laser~\cite{ThorLabsIFO}. The beam is first reflected from a mirror a right angle (to save space on the optical breadboard) and then focused at a converging lens before the $50$/$50$ beamsplitter. The split beam is reflected from mirrors M1 and M2 at the ends of the interferometer arms and recombined at the beamsplitter. The interference pattern is enlarged by a diverging lens after the beamsplitter and projected onto a screen. A commertial speaker is fixed to the back of M2, as the signal input. The interferometer arms lengths are roughly $3\,{\rm cm}$ different and the entire interferometer (without the screen) fits within the dimensions of an A4 piece of paper. We did not include a compensating plate with the beamsplitter, as is often done.} 
	\label{fig:ifo_schematic_webcam}
\end{figure}

The Michelson interferometer configuration used in this work is shown schematically in Figure~\ref{fig:ifo_schematic_webcam}.
It is constructed on a $450 \times 450 \,{\rm mm} $ optical breadboard. 
The laser (shown in the top-left in the figure) has a wavelength of $532\,{\rm nm}$.
In this configuration, the laser beam is first reflected from a mirror which turn the beam $90^{\circ}$ (to save space on the optical breadboard). 
The beam passes through a a converging lens and then to the beamsplitter, which is $50$/$50$ reflection/transmission. 
The beams are reflected by mirrors at the end of each of the arm (\han{$\approx XX\,{\rm cm}$ long}) and are recombined at the beamsplitter. 
After the beamsplitter there is a diverging lens to aid viewing of the interference pattern when it is projected onto a screen as shown.

A photograph of the interference pattern is shown in Figure~\ref{fig:interference_pattern}. 
If the arm lengths were equal, then the interference pattern would be a single dot. 
An intentional offset in the arm lengths of the interferometer produces a pattern with a ring of interference fringes. 
This results in a phase difference between the two beams when they meet at the screen. This path difference is perhaps better seen as the perpendicular separation between one mirror (M1) and the virtual image (M2\textquotesingle) of the other. For an aligned configuration and a fixed separation, the phase difference at a point on the screen only depends on the angle from the centre-line of the beam, and so the pattern is circularly symmetric. When the path difference changes, the fringes stay circular but move radially in or out.
A ring interference pattern is often used in gravitational-wave table-top demonstrations to allow changes in the pattern to be easily seen.

@@

A $0.5\,{\rm W}$ speaker is fixed to the back of one of the mirrors (M2 in Figure~\ref{fig:ifo_schematic_webcam}), providing the audio source for the set-up. 
The speaker is part of a commercial $3.5\,{\rm mm}$-jack, USB powered pair of speakers and was driven by a PC. 
It is attached to the back of the mirror using commercial adhesive putty. 
The other speaker in the pair is kept face-down and as far away from the interferometer as the cables allow to reduce it acting as a second noise source.

@@

Data can be recorded via a webcam or photo-diode. 
In Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} a webcam is used as shown in Figure~\ref{fig:ifo_schematic_webcam}. 
In Section~\ref{sec:optical_microphone}, which considers more complex audio, a photodiode is used as it provides a higher sample rate than the webcam.

>>>>

The basic physics of gravitational-wave detection can be demonstrated with a table-top Michelson interferometer, an optical configuration commonly used in undergraduate labs. 
A simple Michelson interferometer comprises of a laser, a beamsplitter, and two mirrors. 
Laser light is split by a beamsplitter into two arms at right angles to each other. 
At the end of each arm is a mirror which reflects the beams back to the beamsplitter where they recombine. 
The combination of the two beams produces an interference pattern which is dependent on the relative distance traveled by each of the beams. 
A change in arm length results in a change in the interference pattern.

Gravitational-wave detectors are large, complex interferometers.
The LIGO and Virgo detectors have $4\,{\rm km}$ and $3\,{\rm km}$ arms respectively.
They are sensitive to miniscule changes in distance, for example the first detection of a binary black hole merger had a peak gravitational-wave strain of $10^{-21}$ ~\cite{GW150914}.
However, their design is based on the Michelson interferometer. 
Detection and analysis concepts can be demonstrated with table-top optical equipment.

A commonly-used analogy in explaining gravitational-wave science is sound. 
Signals of binary black hole and neutron star mergers can be converted into audio as an aid to explanations~\cite{SoundsOfSpaceTime:online,BlackHoleHunter:online} (however it is important to note that gravitational waves are not sound).  
Sound vibrations provide a simple means to move the components of a demonstration interferometer. 
The vibrations move the mirrors, changing the length of the interferometer arms, and the resulting interference pattern. 
We use sound throughout this work as a means of simulating gravitational-wave-like signals and the data analysis and signal processing techniques used to detect them.

\begin{figure}
% 	\includegraphics[width=0.49\textwidth]{figures/ifo_schematic_webcam.pdf}
	\includegraphics[width=0.49\textwidth]{figures/ifo_schematic_webcam_edit.pdf}
	\caption{\label{fig:ifo_schematic_webcam}
Schematic of the table-top Michelson interferometer. 
The laser (top-left) is a Class 2 $532\,{\rm nm}$ laser~\cite{ThorLabsIFO}. 
The beam is first reflected from a mirror at a right angle and the focused at a converging lens before the beamsplitter which is $50$/$50$ in reflection and transmission. 
The split beams are reflected from mirrors (M1 and M2) at the ends of the interferometer arms and are recombined at the beamsplitter. 
The beam is enlarged by a diverging lens after the beamsplitter. 
The interference pattern is projected onto a screen and recorded using a commercial webcam connected to a personal computer (PC). 
A speaker fixed to the back of M2 provides the signal input from the PC.  
%Michelson interferometer schematic with interference pattern recorded by a commercial webcam. The laser (top-left) is a Class 2 $532\,{\rm nm}$ laser~\cite{ThorLabsIFO}. The beam is first reflected from a mirror a right angle (to save space on the optical breadboard) and then focused at a converging lens before the $50$/$50$ beamsplitter. The split beam is reflected from mirrors M1 and M2 at the ends of the interferometer arms and recombined at the beamsplitter. The interference pattern is enlarged by a diverging lens after the beamsplitter and projected onto a screen. A commertial speaker is fixed to the back of M2, as the signal input. The interferometer arms lengths are roughly $3\,{\rm cm}$ different and the entire interferometer (without the screen) fits within the dimensions of an A4 piece of paper. We did not include a compensating plate with the beamsplitter, as is often done.} 
    }
\end{figure}

The Michelson interferometer configuration used in this work is shown schematically in Fig.~\ref{fig:ifo_schematic_webcam}.
It is constructed on a $450 \times 450 \,{\rm mm} $ optical breadboard. 
The laser (top-left) has a wavelength of $532\,{\rm nm}$.
In the configuration shown in Fig.~\ref{fig:ifo_schematic_webcam} the laser beam is first reflected from a mirror which turns the beam $90^{\circ}$ (to save space on the optical breadboard). 
The beam passes through a a converging lens and then to the beamsplitter, which is $50$/$50$ reflection and transmission. 
The beams are reflected by mirrors at the end of each of the two arms (\han{$\approx XX\,{\rm cm}$ and $\approx XX\,{\rm cm}$ long}) and are recombined at the beamsplitter. 
The resulting interference pattern is projected onto a screen and recorded by a webcam.

The interference pattern is a ring of interference fringes as shown by the photograph in Fig.~\ref{fig:interference_pattern}. 
The pattern is enlarged by a diverging lens between the beamsplitter and the screen to aid in viewing. 
A change in the relative length of the interferometer arms results in the interference rings moving radially inwards or outwards. 
Ring interference patterns like this are often used in gravitational-wave table-top demonstrations to allow changes in the pattern to be easily viewed. 
%This results in a phase difference between the two beams when they meet at the screen. This path difference is perhaps better seen as the perpendicular separation between one mirror (M1) and the virtual image (M2\textquotesingle) of the other. For an aligned configuration and a fixed separation, the phase difference at a point on the screen only depends on the angle from the centre-line of the beam, and so the pattern is circularly symmetric. When the path difference changes, the fringes stay circular but move radially in or out.
%A ring interference pattern is often used in gravitational-wave table-top demonstrations to allow changes in the pattern to be easily seen.

@@

A $0.5\,{\rm W}$ speaker is fixed to the back of one of the mirrors (M2 in Fig.~\ref{fig:ifo_schematic_webcam}), providing the audio source for the set-up. 
The speaker is part of a commercial $3.5\,{\rm mm}$-jack, USB powered pair of speakers and was driven by a PC. 
It is attached to the back of the mirror using commercial adhesive putty. 
The other speaker in the pair is kept face-down and as far away from the interferometer as the cables allow to reduce it acting as a second noise source.

@@

Data can be recorded via a webcam (as shown in Fig.~\ref{fig:ifo_schematic_webcam}) or photo-diode. 
In Sections~\ref{sec:single_tone} and~\ref{sec:viterbi_wandering} a webcam is used. 
In Section~\ref{sec:optical_microphone}, which considers more complex audio, a photodiode is used as it provides a higher sample rate than the webcam.
